Tampering Detection In Low-Power \\ Smart Cameras

### 2015-06-20 paper.tex
Sections 1-3

\begin{abstract}
A desirable feature in smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene. No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., occlusions or the device displacement), these have to be promptly detected to possibly activate countermeasures. Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at full-speed  frame-rates, nor use sophisticated image-analysis algorithms.

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous indicator is detected as a tampered frame. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to improve the detection of camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, smart cameras, displacement detection, blurring detection}
\end{abstract}

\section{Introduction}\label{sec:introduction}
% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Figure \ref{fig:neve}. Similarly, intentional attacks like displacing the camera Figure \ref{fig:pioggia} \gi{Adriano SISTEMA CON LE NUOVE FIGURE}, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes. We refer to these events/attacks as tampering. In many situations, tampering is not straightforward to detect (e.g. when the camera is not physically damaged and does not go out-of-order), and image analysis techniques are the only viable option.

\begin{figure}[t!]
\centering
\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.3\linewidth]{Immagini/pioggia}}
\subfigure[]{\label{fig:neve} \includegraphics[width=0.3\linewidth]{Immagini/neve}}
\caption[Tampering examples]{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} Blur due to rain drops on the camera lens. \textbf{(b)} Occlusion due to snow on the camera lens.}
\label{fig:tampering}
\end{figure}

% Why is it interesting and important?
Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts. In fact, tampering prevents the correct interpretation of the scene: for instance, even a mild blur might hinder the identification of important details such as licence plates. Surveillance cameras are typically connected to the power supply and operate at normal frame-rates (e.g. around few frames per second): in these conditions, several tampering-detection algorithms have been presented. In particular, camera displacement and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop. In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain. Two background models, estimated over different time intervals, are used in~\cite{saglam2009real}. In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation. Background estimates and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while the number of SURF~\cite{bay2006surf} keypoints is used to detect blurring. A frame buffer rather than an explicit background is used~\cite{ribnick2006real}. All these above solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.

In this work we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and produced by STMicroelectronics. SecSoC is a battery-powered device, characterized by a constrained computational power (clock rates 82.5 MHz at 1.2V, and sub-1MHz at 0.6V) and reduced memory (1.25 MB). While low-power cameras are not meant for critical surveillance applications, they might be easily employed in distributed systems for monitoring wide environments thanks to their low cost and maintenance requirements.  %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests. 
% These units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting~\cite{magno2009adaptive}. 
%In practice, because of energy constrain, low-power devices have to operate at very low frame-rates: possibly less than one frame every minute.

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power cameras (eventually organized in wireless sensor network) is more challenging than in conventional surveillance systems, and this problem has not been much investigated so far~\cite{perrig2004security,alippi2010detecting}. Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly. These aspects prevent the use of learned background models and the analysis of foreground variations. In fact, in dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions (see Figure~\ref{fig:sequences}). Smart cameras have then to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames. %Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and unnecessary network transmissions.
% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose. Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).

%What are the key components of my approach and results? Also include any specific limitations.
We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}. The proposed algorithm (Section~\ref{sec:propSol}) relies on indicators that can be easily computed in low-power smart cameras, such as the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and detects tampering as frames yielding outliers. In particular, we show that separately monitoring indicators over different regions of the image can substantially improve the displacement-detection performance. Remarkably, image regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t. monitoring the whole image (Section \ref{subsec:coputationalComplexity}). Thus, our solution represents a prompt trigger that can be possibly combined with other sequential monitoring techniques. Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance. % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

%\subsection{Related Works}\label{subsec:relWorks}
%% The literature concerning tampering detection is mostly focused on video surveillance applications and process at few frames per second.
% 
%%
%: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
%\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
%
%
%
\section{Problem Formulation}\label{sec:probForm}
%
Let $z_t$ be the frame acquired at time $t$
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes an operator that transforms the original image $y_t$ in the acquired frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates. As far as there are no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though $y_t$ might show different content because of scene changes.\\
When, at time $\tau^*$, an external disturbance introduces \emph{blurring}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
\begin{equation}
\label{eq:model_defocus}
\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \geq \tau^*
\end{equation}
where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.\\
A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
\begin{equation}
\label{eq:model_displacement}
z_t(x)  = \left\{ \begin{array}{rcl}
y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
\end{array}\right. ,
\end{equation}
where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/sequenze}
\caption{Sequences taken from webcams: \textbf{(a)} defocus event on 3-rd frame, created using gaussian filtering; \textbf{(b)} displacement event on 3-rd frame, created moving the crop of the original image from the center.}
\label{fig:sequences}
\end{figure}

@@

Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$). The first two are meant to detect camera displacement, while the latter image blurring, as this is expected to cause a drop in the high-frequency content. As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$. The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.

For each $z_t$, the luma is separately computed within each of the $K$ regions:
\begin{equation}\label{eq:lumaRegions}
l_k(t) =\frac{1}{\#R_k} \sum_{x \in R_k} z_t(x), \ \ k = 1, \dots, K\,,
\end{equation}
and, similarly, we define the average frame difference
\begin{equation}\label{eq:FDRegions}
d_k(t) =\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,
\end{equation}
while the average gradient norm is computed over the whole image
\begin{equation}
\label{eq:normaGradiente}
g(t) = \sum_{x \in X} \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
\end{equation}
where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, $\circledast$ denotes the 2d convolution and $\#(\cdot)$ the cardinality of a set. The choice of using luma or frame difference also depends on the memory availability to store the frame. In what follows and in Algorithm~\ref{alg:DISPL} we will consider only the luma $l$ even though the same applies to frame difference $d$.

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{Immagini/indicatori}
\caption{Example of camera displacement. Left plots report the luma values, while right plots depict the detrended luma: we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}). Red stars indicate a single displaced frame, which occurs at $t =157$. Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected. }
\label{fig:indicatori} 
\end{figure}
%Tampering changes the sequence of indicators:
Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the luma. First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)). Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here. In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here indicators follow an unpredictable trend because of changes in the scene or in the illumination. Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative (line~\ref{DISPL-Test5})
\begin{equation}\label{eq:detrending}
 \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
\end{equation}
and we similarly define $\overline{\partial d}_k(t)$ and $\partial g(t)$.

@@

	\textbf{Operational phase}:\\
	\lnl{DISPL-Test1} \For{$t=T_{o} + 1,\dots,\infty$}{
		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$;\\
		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
			\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
					\lnl{DISPL-Test9} tampering (probably blurring) has occurred in $z_t$ \\
									}
		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
			\lnl{DISPL-Test5} Compute $\partial l_k(t)$\\
			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma_l \sigma_{l_k} \vee \partial l_k(t) > \gamma_l \sigma_{l_k} $}{
				\lnl{DISPL-Test7} $n_l=n_l+1$\\
			}
		}
		\lnl{DISPL-Test8} \If{$n_l\geq \Gamma_l$}{
			\lnl{DISPL-Test9} tampering (probably displacement or occlusion) has occurred in $z_t$ \\
		}
	}   
	\caption{The Proposed Tampering-Detection Algorithm}
	\label{alg:DISPL}
\end{algorithm}

The sequences of detrended indicators (reported in Figure~\ref{fig:indicatori}) can be suitably monitored by the following confidence intervals:
\begin{equation}\label{eq:confidenceRegions}
 [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}], \ \  k = 1,\dots,K\,,
\end{equation}
where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (lines \ref{DISPL-Tr1} and \ref{DISPL-Tr7}), computed from tampering-free frames provided for training (i.e. $z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter. A similar interval is built for  $\overline{\partial d}_k(t)$ and $\partial g$ (line~\ref{DISPL-Tr7}).

During operations, any indicator falling outside its confidence region is considered an outlier (line~\ref{DISPL-Test6}): any outlier in $\partial g$ yields a blurring alert (line~\ref{DISPL-Test11}). Camera displacements alerts are raised when at least $\Gamma_l$ indicators $\partial l_k$  simultaneously yield outliers (line~\ref{DISPL-Test8}). The threshold $\Gamma_l$ together with $\gamma_l$ determine the displacement-detection promptness. The extreme cases are $\Gamma_l = 1$, where a displacement alert raises at any region firing an outlier, and $\Gamma_l = K-1$ where all but one region have to simultaneously fire an outlier (it is worth excluding a region since, for instance the sky-region typically does not change when the camera is displaced, either horizontally or vertically).

It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yield two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of few consecutive tampered frames yield an outlier in the detrended indicators. This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame. On the one hand, this one-shot solution can provide prompt detections, and this is particularly important at the low frame-rates we consider. On the other hand, one-shot monitoring disregards the fact that tampering might persist over several frames. To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, and possibly combined with the proposed one-shot monitoring.
%
%
\subsection{Scene Segmentation}\label{subsec:Segmentation}
Scene has to be preliminarily segmented to define regions that our algorithm takes as input (see Algorithm \ref{alg:DISPL}). To this purpose, we use part of the training frames before $T_0$ to compute a feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$ 
\begin{equation}
\label{eq:featureVector}
\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X\,.
\end{equation}
In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed overtime. These feature vectors are meant to cluster pixels in regions having similar spatial appearance and temporal behavior over training frames.

>>>>

\begin{abstract}
A desirable feature in smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene. No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., occlusions or device displacements), these have to be promptly detected to possibly activate countermeasures. Tampering detection is particularly challenging in battery-powered cameras, where it is not possible to acquire images at full-speed frame-rates, nor use sophisticated image-analysis algorithms.

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras. The algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme: any frame yielding an outlier is detected as tampered. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to improve the detection of camera displacements. Experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, smart cameras, displacement detection, blurring detection}
\end{abstract}

\section{Introduction}\label{sec:introduction}
% What is the problem?
Cameras operating outdoor and in harsh environments are exposed to atmospheric phenomena and intentional attacks that might prevent the correct image acquisition. Rain, snow, dust lying on the camera lens cause blurry (Figure \ref{fig:pioggia}) or partially occluded (Figure \ref{fig:neve}) pictures, while wind might displace the camera (Figure \ref{fig:displ1} -\ref{fig:displ2}). Similarly, an attacker can intentionally change the camera focus, spray some opaque or glossy liquid, displace or occlude the camera. We refer to these events/attacks as tampering. In many situations, tampering is not straightforward to detect (e.g. when the camera is not physically damaged and does not go out-of-order), and image analysis techniques are the only viable option.

\begin{figure}[t!]
\centering
\subfigure[]{\label{fig:pioggia} \includegraphics[width=0.225\linewidth]{Immagini/pioggia}}
\subfigure[]{\label{fig:neve} \includegraphics[width=0.225\linewidth]{Immagini/neve}}
\subfigure[]{\label{fig:displ1} \includegraphics[width=0.225\linewidth]{Immagini/displ1}}
\subfigure[]{\label{fig:displ2} \includegraphics[width=0.225\linewidth]{Immagini/displ2}}
\caption{Examples of tampering events due to atmospheric phenomena. \textbf{(a)} rain drops resulting in a blurry picture, \textbf{(b)} snow partially occluding the camera view, \textbf{(c)}-\textbf{(d)} wind displacing the camera.}
\label{fig:tampering}
\end{figure}

% Why is it interesting and important?
Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, which are expected to autonomously detect any tampering, and promptly report alerts. Tampering, in fact, might result in images that are useless for monitoring purposes: even a mild blur might hinder the identification of important details such as licence plates. Surveillance cameras most often operate at normal frame-rates (e.g. around few frames per second) and are connected to the power supply. Tampering-detection for surveillance cameras have been quite investigated in the literature; in particular, camera displacements and occlusions are typically detected by comparing the current frame against an estimate of the scene background, while blurring is detected by monitoring the high-frequency components of images, as these are expected to drop. In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing the histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain. Two background models, estimated over different time intervals, are used in~\cite{saglam2009real}. In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation. Background estimates and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while the number of SURF~\cite{bay2006surf} keypoints is used to detect blurring. A buffer of recent frames rather than an explicit background is used~\cite{ribnick2006real}. All the above solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.

In this work we expressly target low-power and ultra-low-power smart cameras, like \emph{SecSoC} (Security System on Chip), an innovative prototype based on a cluster of ReISC (Reduced Energy Instruction Set Computer) cores, designed and produced by STMicroelectronics. SecSoC is a battery-powered device, characterized by a constrained computational power (clock rates 82.5 MHz at 1.2V, and sub-1MHz at 0.6V) and reduced memory (1.25 MB). While low-power cameras are not meant for critical surveillance applications, they might be easily employed for monitoring wide environments thanks to their low cost and maintenance requirements.  %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests. 
% These units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting~\cite{magno2009adaptive}. 
%In practice, because of energy constrain, low-power devices have to operate at very low frame-rates: possibly less than one frame every minute.

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power cameras (eventually organized in wireless sensor network) is more challenging than in conventional surveillance systems, and this problem has not been much investigated so far~\cite{perrig2004security,alippi2010detecting}. Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly. These aspects prevent the use of learned background models and the analysis of foreground variations. In fact, when dynamic environments are monitored at low frame-rates, changes in the scene and in the light conditions might produce consecutive frames that are very different (see Figure~\ref{fig:sequences}). Smart cameras have to promptly distinguish between \emph{normal} changes (due to illumination or movements in the scene), and changes due to camera tampering, to raise an alert and eventually avoid the transmission of tampered frames. %Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and unnecessary network transmissions.
% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose. Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).

%What are the key components of my approach and results? Also include any specific limitations.
We address the detection of camera blurring and displacement (Section \ref{sec:probForm}), and propose an algorithm (Section~\ref{sec:propSol}) that relies on indicators that can be easily computed in low-power smart cameras. We monitor the average image intensity or frame difference (to detect camera displacement), and the average gradient norm (to detect blurring), and consider any frame yielding an outlier as tampered. In particular, we show that separately monitoring these indicators over different regions of the image can substantially improve the displacement-detection performance. Image regions are defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t. monitoring the whole image (Section \ref{subsec:coputationalComplexity}). Our algorithm thus represents a prompt trigger, to be possibly combined with other sequential monitoring techniques. Experiments (Section \ref{sec:experiments}) show that leveraging image regions can substantially improve the displacement-detection performance. % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

%\subsection{Related Works}\label{subsec:relWorks}
%% The literature concerning tampering detection is mostly focused on video surveillance applications and process at few frames per second.
% 
%%
%: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
%\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
%
%
%
\section{Problem Formulation}\label{sec:probForm}
%
Let $z_t$ be the frame acquired at time $t$
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x), \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes an operator transforming the original image $y_t$, and $x\in \mathbb{Z}^2$ indicates the pixel coordinates belonging to the regular pixel grid $X \subset \mathbb{Z}^2$. As far as there are no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x), \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for image noise and $y_t$ are acquired from the same viewpoint and camera orientation, even though typically $y_t \neq y_{t-1}$ because the depicted scene changes.\\
When, at time $\tau^*$, an external disturbance introduces \emph{blurring}, the image $y_t$ is degraded by an unknown blur operator, and $z_t$ becomes
\begin{equation}
\label{eq:model_defocus}
\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \geq \tau^*
\end{equation}
where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.\\
A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
\begin{equation}
\label{eq:model_displacement}
z_t(x)  = \left\{ \begin{array}{rcl}
y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
\end{array}\right. ,
\end{equation}
where $w_t$ and $y_t$ refer to different viewpoints and/or camera orientations.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/sequenze}
\caption{Examples of blurring and displacement synthetically introduced: \textbf{(a)} the 3-rd frame is blurred by a convolution against a Gaussian filter; \textbf{(b)} the 3-rd frame has been displaced by cropping a different region in the original image.}
\label{fig:sequences}
\end{figure}

@@

Algorithm~\ref{alg:DISPL} presents the proposed tampering detection, which relies on simple indicators such as the average intensity (the \emph{luma}, denoted by $l$), the average frame difference (denoted by $d$) and the average norm of the gradient (denoted by $g$). The first two are meant to detect camera displacements, which would substantially change the image content, while the latter the blurring, which would cause a drop in the high-frequency content. As anticipated in Section~\ref{sec:introduction}, during the initial configuration the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$. The employed segmentation is detailed in Section~\ref{subsec:Segmentation}.

For each $z_t$, we compute the luma and frame difference separately on each of the $K$ regions,
\begin{eqnarray}
\vspace{-0.5cm}
l_k(t) &=& \frac{1}{\#R_k} \sum_{x \in R_k} z_t(x) \label{eq:lumaRegions}\\
d_k(t) &=&\frac{1}{\#R_k} \sum_{x \in R_k} \left(z_t(x) - z_{t-1}(x) \right)^2, \ \ k = 1, \dots, K\,,\label{eq:FDRegions}
\vspace{-0.5cm}
\end{eqnarray}
where $\#(\cdot)$ denotes the cardinality of a set. We can simultaneously monitor both luma and frame difference, even though computing $d_k$ requires to store the previous frame and this also depends on memory availability of the device. In what follows, including Algorithm~\ref{alg:DISPL}, we consider only the luma $l$ but the same procedures apply to the frame difference $d$ as well. The average gradient norm is instead computed over the whole image
\begin{equation}
\label{eq:normaGradiente}
g(t) = \sum_{x \in X} \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
\end{equation}
where $f_h$ and $f_v$ are the horizontal and vertical derivative filters, respectively, and $\circledast$ denotes the 2d convolution.

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{Immagini/indicatori}
\caption{Example of camera displacement. Left plots report the luma values, while right plots depict the detrended luma: we show both indicators computed over the full image (\textbf{a}) and on two reported regions (\textbf{b} and \textbf{c}). Red stars indicate a single displaced frame, which occurs at $t =157$. Displacement yields a peak in $l$ and two outliers in the sequence of detrended indicators $\partial l$, which can be clearly detected. }
\label{fig:indicatori} 
\end{figure}
%Tampering changes the sequence of indicators:
Figure~\ref{fig:indicatori}(a) shows how a camera displacement affects the luma. First of all, we observe that a displaced frame changes $l(t)$ (in this case introducing a peak) and that this change is also visible in the indicators $l_k(t)$ for the regions (Figure~\ref{fig:indicatori}(b) and Figure~\ref{fig:indicatori}(c)). Second, we observe that outlier-detection methods~\cite{Chandola2009} based on density estimates or on confidence intervals --that are very efficient to run-- cannot be straightforwardly applied here. In fact, these methods are meant for independent and identically distributed (i.i.d.) random variables, while here indicators follow an unpredictable trend because of changes in the scene or in the illumination. Therefore, we perform a detrending~\cite{Gustafsson2000} of the indicator sequence by a temporal derivative
\begin{equation}\label{eq:detrending}
 \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
\end{equation}
and we similarly define $\partial d_k(t)$ and $\partial g(t)$.

@@

	\textbf{Operational phase}:\\
	\lnl{DISPL-Test1} \For{$t=T_{o} + 1,\dots,\infty$}{
		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$;\\
		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
			\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma_g \sigma_{g} \vee \partial l(t) > \gamma_g \sigma_{g} $}{
					\lnl{DISPL-Test9} raise an alert: tampering (probably blurring) in $z_t$ \\
									}
		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
			\lnl{DISPL-Test5} Compute $\partial l_k(t)$\\
			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma_l \sigma_{l_k} \vee \partial l_k(t) > \gamma_l \sigma_{l_k} $}{
				\lnl{DISPL-Test7} $n_l=n_l+1$\\
			}
		}
		\lnl{DISPL-Test8} \If{$n_l\geq \Gamma_l$}{
			\lnl{DISPL-Test9} raise an alert: tampering (probably displacement or occlusion) in $z_t$ \\
		}
	}   
	\caption{The Proposed Tampering-Detection Algorithm}
	\label{alg:DISPL}
\end{algorithm}

The sequences of detrended indicators (reported in Figure~\ref{fig:indicatori}) can be suitably monitored by the following confidence intervals:
\begin{equation}\label{eq:confidenceRegions}
 [\overline{\partial l}_k(t) - \gamma_l \sigma_{l_k}, \overline{\partial l}_k(t) + \gamma_l \sigma_{l_k}], \ \  k = 1,\dots,K\,,
\end{equation}
where $\overline{\partial l}_k(t)$ denotes the mean and $\sigma_{l_k}$ the standard deviation of $\partial l$ over $R_k$ (Algorithm~\ref{alg:DISPL}, line \ref{DISPL-Tr1}), computed from tampering-free frames provided for training (i.e. $z_t\,, t = 1, \dots, T_o$) and $\gamma_l >0$ is a tuning parameter. Similar intervals are built for  $\partial d_k(t)$ and $\partial g$ (line~\ref{DISPL-Tr7}).

During operations, indicators are computed (lines~\ref{DISPL-Test10} and~\ref{DISPL-Test5}), and any indicator falling outside its confidence region is considered an outlier. In particular, any outlier in $\partial g$ yields a blurring alert (line~\ref{DISPL-Test11}), while camera-displacement alerts are raised when at least $\Gamma_l$ indicators $\partial l_k$  simultaneously yield outliers (line~\ref{DISPL-Test8}). The threshold $\Gamma_l$ together with $\gamma_l$ determine the displacement-detection promptness. The extreme cases are $\Gamma_l = 1$, where it is enough that a single region fires an outlier to raise a camera-displacement alert, and $\Gamma_l = K-1$ where all but one region\footnote{It is better to exclude a region since, for instance, the sky-region typically does not change when the camera is displaced, either horizontally or vertically} have to simultaneously fire an outlier to raise the alert.

It is important to remark that tampering yields outliers in the transient of the detrended indicators~\eqref{eq:detrending}: namely, a single tampered frame yields two outliers (as shown in Figure~\ref{fig:indicatori}) and only the first and the last of a sequence of consecutive tampered frames yield outliers in the detrended indicators. This is the reason why detrended indicators are monitored in a one-shot manner, targeting the detection of the first tampered-frame. On the one hand, this one-shot solution can provide prompt detections, which is particularly important at the low frame-rates we consider, on the other hand, the persistence of tampering is disregarded. To take this valuable information into account, some form of sequential monitoring should be applied to the indicator sequence as in~\cite{alippi2010detecting}, possibly combined with the proposed one-shot monitoring.
%
%
\subsection{Scene Segmentation}\label{subsec:Segmentation}
Scene has to be preliminarily segmented to define regions that Algorithm \ref{alg:DISPL} takes as input. To this purpose, we use part of the training frames before $T_0$ to compute the feature vector $\textbf{f}(x)\in \mathbb{R}^5$ for each pixel $x\in\mathcal{X}$ 
\begin{equation}
\label{eq:featureVector}
\textbf{f}(x)=\left[r(x);c(x);\bar{l}(x);\sigma_{l}(x);\overline{g}(x);\sigma_{g}(x)\right]\,, \ \forall x \in X\,.
\end{equation}
In \eqref{eq:featureVector}, $r(x)$ and $c(x)$ denotes the row and column of $x$, respectively, $\bar{l}(x)$ and $\sigma_{l}(x)$ the mean and standard deviation of the intensity at $x$, computed over time, and $\overline{g}(x)$ and $\sigma_g(x)$ the mean of gradient norm and its standard deviation, computed overtime. These feature vectors are meant to cluster pixels in regions having, over training frames, similar spatial appearance and temporal behavior.
